Dynamical Properties of Molecular Cloud Complexes in Galaxies at the Epoch of Reionization

### 2018-10-25 WRITEUP/ms.tex
organization, shorten abstract

\citestyle{aa}
\shorttitle{Dynamical Properties of Molecular Cloud Complexes in $z$\ssim6 Prototypical Galaxy
}
\shortauthors{Leung et al.}

\begin{document}
\title{
Dynamical Properties of Molecular Cloud Complexes in a Redshift 6 Prototypical Galaxy at the Epoch of Reionization
}

\author{T. K. Daisy Leung\altaffilmark{1, 2}}
\author{Andrea Pallottini\altaffilmark{3, 4}}
\author{Andrea Ferrara\altaffilmark{4, 5}}
\author{Mordecai-Mark Mac Low\altaffilmark{2, 6, 7
%mm, 8   [I only use the Drexel affiliation when working with folks
%from there...]
}}

\affil{\textsuperscript{1} Department of Astronomy, Space Sciences Building, Cornell University, Ithaca, NY 14853, USA; }
\email{tleung@astro.cornell.edu}
\altaffiltext{2}{Center for Computational Astrophysics, Flatiron Institute, 162 Fifth Avenue, New York, NY 10010, USA}
\altaffiltext{3}{Centro Fermi, Museo Storico della Fisica e Centro Studi e Ricerche ``Enrico Fermi'', Piazza del Viminale 1, Roma, 00184, Italy}
\altaffiltext{4}{Scuola Normale Superiore, Piazza dei Cavalieri 7, I-56126 Pisa, Italy}
\altaffiltext{5}{Kavli Institute for the Physics and Mathematics of the Universe (IPMU), The University of Tokyo, 5-1-5 Kashiwanoha, Kashiwa 277-8583, Japan}
\altaffiltext{6}{Institut f{\"u}r Theoretische Astrophysik, Zentrum f{\"u}r Astronomie der Universit{\"a}t Heidelberg, 69120 Heidelberg, Germany}
\altaffiltext{7}{American Museum of Natural History, 79th St.~at Central Park West, New York, NY 10024, USA}
%mm \altaffiltext{8}{Department of Physics, Disque Hall, Drexel University, Philadelphia, PA 19104, USA}

\begin{abstract}
We study the dynamical properties of molecular clouds complexes (MC) and their temporal evolution in \flower, a prototypical galaxy
at the Epoch of Reionization (EoR) using cosmological zoom-in simulation.
We use a clump-finding algorithm and a set of H$_2$ volumetric densities ($n_{\rm cut}$) 
to identify MCs and their sub-structures.
We extract properties such as mass, size, velocity dispersion, gas surface density, and virial parameter ($\alpha_{\rm vir}$) 
for each MC and 
compare them with those observed in the Milky Way disk, the Galactic center,
and gas-rich starburst galaxies in the local Universe and at the peak epoch of cosmic \SF.
The high velocity dispersions found in the MCs of \flower ($\sigma\approx$\,200\,\kms) are comparable to 
those observed in starburst galaxies.
The mass scale of the MCs is on the order of $10^{6.5-9}$\,\Msun. The $\sim$200\,pc-scale MCs found with a low density threshold 
correspond to the spiral arms of \flower which break down into smaller $\lesssim$100\,pc-scale MCs at higher density thresholds. 
The more massive and bigger MCs in \flower compared to the Milky Way 
likely result from the higher gas mass fraction, surface density, and velocity dispersion, 
which set the scale for fragmentation.
The MCs of \flower are found to have higher $\sigma$ and $\Sigma$ systematically regardless of the $n_{\rm cut}$ adopted,
with $\sigma\gtrsim$100\,\kms even when we increase the $n_{\rm cut}$ and even for  
the molecular substructures.
The high velocity dispersions likely result from 
the strong supernova and stellar feedback \flower experienced over the multiple episodes of bursty \SF.
Virial analysis indicates that the MCs found in the main disk of \flower are unbound, but its substructures have lower virial parameters. 
This is consistent with the notion that collapsing structures result from gravitational instability within globally stable structures, which are 
supported by turbulence and rotation on large scale.
We find $\alpha_{\rm vir}\approx$\,2 for the MCs in the satellite galaxies, which we interpret as result of the weaker stellar feedback 
due to the less frequency \SF compared to \flower (also supported by the lower stellar-to-gas mass ratio of the former).
MCs in the satellites are therefore likely collapsing structures. This paints a picture, in which at the EoR, \SF continues 
as gas is being accreted from the intergalactic medium.
We find no temporal variations in the MC dynamics over the course of 700\,Myr traced in our simulation, at least in terms of the scaling relations examined. 
Our results are insensitive to the choice of $n_{\rm cut}$, except for the slope of the cumulative 
mass distribution, which steepens with $n_{\rm cut}$. 
High resolution imaging of the first galaxies with the Atacama Large (sub-)Millimeter Array (ALMA) and the 
Next Generation Very Large Array (ngVLA)
will provide useful observational data to test our findings and the validity of our simulation to shed
light on \SF since the cosmic dark ages.
\end{abstract}
\keywords{methods: data analysis --
          galaxies: high-redshift --
          galaxies: ISM --
          galaxies: evolution --
          galaxies: formation --
          galaxies: starburst --
          stars: formation}

@@

While ALMA has enabled the detections of
ISM diagnostic lines e.g., the \cii158\,$\micron$ and CO line emission in
normal (SFR$<$\,100\,\Msun\,yr\pmOne) galaxies at \z$>$\,6 over the past few years \citep[e.g.,][]{Odorico18a, Carniani18b},
we are still far from mapping their molecular ISM due to the cosmological dimming.
As such, we have undertaken a study, exploiting
state-of-the-art cosmological zoom-in hydrodynamic simulation
\ncode{Serra} (Greenhouse in Italian; \citealt{Pallottini17a, Pallottini17b}), to examine
the dynamical properties of the molecular cloud complexes in a \z$\sim$6 prototypical (i.e., $L^*$) 
galaxy.

@@

% interpretation
Velocity dispersion and surface density are tightly connected to pressure. 
That is, the pressure of an MC is expected to increase with its velocity dispersion and surface density (see e.g., \Eq{pext} and 
since $P\propto\Sigma^2$).
In the central molecular zone of the Milky Way (CMZ), we have seen molecular clouds with high pressures of 
$P/k_B > $10$^7$\,K\,cm$^{-3}$. Such high pressures are also seen in massive star-forming regions such as 30 Doradus.
The higher $\Sigma_{\rm gas}$ in turn is related to more massive clouds being formed.
In M64 and M51, there are a few clouds with masses $M_{\rm cl}>$10$^7$\,\Msun \citep{Rosolowsky05a, Koda09a, Colombo14a}. 
On the other hand, such massive molecular structures are unknown to Local Group galaxies 
\citep[e.g.,][]{Rice16a}, and are likely the sites 
from which massive star clusters or OB associations are to be formed.
% 
The shallower slope found in the CMF for \flower compared to those observed in the nearby Universe (see \Sec{cmf}) may
suggest that the MCs of \flower are more massive compared to those observed in the nearby Universe,
notwithstanding the limited numerical resolution of our simulation.
That said, the Jeans mass $M_J$\eq$\sigma^4/(G^2\Sigma)$ yields $\sim$\,8\E{8}\,\Msun 
for a velocity dispersion of $\sigma\approx50$\,\kms and a gas mass surface density of 
$\Sigma_{\rm gas}\approx450$\,\Msun\,pc$^{-2}$, which are typical of the MCs of \flower.
This Jeans mass is in agreement with the mass range found (see \Fig{dist}).

>>>>

\citestyle{aa}
\shorttitle{Dynamical Properties of Molecular Cloud Complexes in Galaxies at the Epoch of Reionization
}
\shortauthors{Leung et al.}

\begin{document}
\title{
Dynamical Properties of Molecular Cloud Complexes in Galaxies at the Epoch of Reionization
}

\author{T. K. Daisy Leung\altaffilmark{1, 2}}
\author{Andrea Pallottini\altaffilmark{3, 4}}
\author{Andrea Ferrara\altaffilmark{4, 5}}
\author{Mordecai-Mark Mac Low\altaffilmark{2, 6, 7}}

\affil{\textsuperscript{1} Department of Astronomy, Space Sciences Building, Cornell University, Ithaca, NY 14853, USA; }
\email{tleung@astro.cornell.edu}
\altaffiltext{2}{Center for Computational Astrophysics, Flatiron Institute, 162 Fifth Avenue, New York, NY 10010, USA}
\altaffiltext{3}{Centro Fermi, Museo Storico della Fisica e Centro Studi e Ricerche ``Enrico Fermi'', Piazza del Viminale 1, Roma, 00184, Italy}
\altaffiltext{4}{Scuola Normale Superiore, Piazza dei Cavalieri 7, I-56126 Pisa, Italy}
\altaffiltext{5}{Kavli Institute for the Physics and Mathematics of the Universe (IPMU), The University of Tokyo, 5-1-5 Kashiwanoha, Kashiwa 277-8583, Japan}
\altaffiltext{6}{Institut f{\"u}r Theoretische Astrophysik, Zentrum f{\"u}r Astronomie der Universit{\"a}t Heidelberg, 69120 Heidelberg, Germany}
\altaffiltext{7}{American Museum of Natural History, 79th St.~at Central Park West, New York, NY 10024, USA}

\begin{abstract}
We study the dynamical properties of molecular clouds complexes (MC) and their temporal evolution in a prototypical galaxy (\flower)
and its satellite galaxies at the Epoch of Reionization (EoR) using cosmological zoom-in simulation.
We use a clump-finding algorithm and H$_2$ density to identify MCs and their sub-structures.
We compare the mass, size, velocity dispersion, gas surface density, and virial parameter ($\alpha_{\rm vir}$)
of each MC with those observed in the Milky Way, the Galactic center,
and gas-rich starburst galaxies in the local Universe and at the peak epoch of cosmic \SF.
The high velocity dispersions found in the MCs of \flower ($\sigma\approx$\,200\,\kms) are comparable to
those observed in starburst galaxies.
The mass scale of the MCs is on the order of $10^{6.5-9}$\,\Msun.
The $\sim$200\,pc-scale MCs found with a low density threshold
correspond to the spiral arms of \flower which break down into smaller $\lesssim$100\,pc-scale MCs at higher density thresholds.
The velocity dispersion and gas surface density are found to be systematically higher than nearby galaxies
regardless of the $n_{\rm cut}$ adopted, with $\sigma\gtrsim$\,100\,\kms.
This likely results from the strong supernova and stellar feedback \flower experienced over multiple episodes of
bursty \SF and constant cosmic gas inflows.
Our results suggest that the more massive and bigger MCs of \flower compared to the Milky Way
likely result from the higher gas mass fraction, surface density, and velocity dispersion,
which together sets the scale for fragmentation.
Virial analysis indicates that the MCs found in the main disk of \flower are unbound, but its substructures have lower virial parameters,
 consistent with the notion that collapsing structures result from gravitational instability within globally stable structures, which are
supported by turbulence and rotation on large scale.
We find $\alpha_{\rm vir}\approx$\,2 for the MCs in the satellite galaxies, which we interpret as result of the weaker stellar feedback
due to the less frequency \SF compared to \flower (also supported by the lower stellar-to-gas mass ratio of the former).
High resolution imaging of the first galaxies with the Atacama Large (sub-)Millimeter Array (ALMA) and the
Next Generation Very Large Array (ngVLA)
will provide useful observational data to test our findings and the validity of our simulation to shed
light on \SF since the cosmic dark ages.
\end{abstract}
\keywords{methods: data analysis --
          galaxies: high-redshift --
          galaxies: ISM --
          galaxies: evolution --
          galaxies: formation --
          galaxies: starburst --
          stars: formation}

@@

While ALMA has enabled the detections of
ISM diagnostic lines e.g., the \cii158\,$\micron$ and CO line emission in
normal (SFR$<$\,100\,\Msun\,yr\pmOne) galaxies at \z$>$\,6 over the past few years \citep[e.g.,][]{Odorico18a, Carniani18b},
we are still far from mapping their molecular ISM due to the cosmological dimming.
As such, we have undertaken a study, exploiting
state-of-the-art cosmological zoom-in hydrodynamic simulation
\ncode{Serra} (Greenhouse in Italian; \citealt{Pallottini17a, Pallottini17b}), to examine
the dynamical properties of the molecular cloud complexes in a \z$\sim$6 prototypical (i.e., $L^*$)
galaxy.
{\bf AF: need aim and objectives of the study}

@@

% interpretation
Velocity dispersion and surface density are tightly connected to pressure.
That is, the pressure of an MC is expected to increase with its velocity dispersion and surface density (see e.g., \Eq{pext} and
since $P\propto\Sigma^2$).
In the central molecular zone of the Milky Way (CMZ), we have seen molecular clouds with high pressures of
$P/k_B > $10$^7$\,K\,cm$^{-3}$. Such high pressures are also seen in massive star-forming regions such as 30 Doradus.
The higher $\Sigma_{\rm gas}$ in turn is related to more massive clouds being formed.
In M64 and M51, there are a few clouds with masses $M_{\rm cl}>$10$^7$\,\Msun \citep{Rosolowsky05a, Koda09a, Colombo14a}.
On the other hand, such massive molecular structures are unknown to Local Group galaxies
\citep[e.g.,][]{Rice16a}, and are likely the sites
from which massive star clusters or OB associations are to be formed.
%
The shallower slope found in the CMF for \flower compared to those observed in the nearby Universe (see \Sec{cmf}) may
suggest that the MCs of \flower are more massive compared to those observed in the nearby Universe,
notwithstanding the limited numerical resolution of our simulation.
That said, based on the scaling of $M_J$\eq$\sigma^4/(G^2\Sigma)$ for the Jeans mass,
a velocity dispersion of $\sigma\approx50$\,\kms and a gas mass surface density of
$\Sigma_{\rm gas}\approx450$\,\Msun\,pc$^{-2}$
yield $M_J\sim$\,8\E{8}\,\Msun, which is in good agreement with the mass range found in the MCs identified (see \Fig{dist}).
The enhanced MCs size-scale compared to local GMCs
is therefore likely a manifestation of gravitational instabilities (resulting from the higher $f_{\rm gas}), which drives
supersonic turbulence in the ISM.
